```{r}
```{r}
# preprocessing

# 01. LOAD LIBRARIES 
# Load all necessary libraries for data wrangling, modeling, and visualization
library(tidyverse)
library(caret)
library(ranger)
library(pROC)
library(FactoMineR)
library(factoextra)
library(DMwR)
library(xgboost)
library(survival)
library(survminer)
library(pheatmap)
library(ggrepel)
library(ComplexUpset)


# ===================== 02. LOAD & CLEAN DATA =====================
# Load METABRIC data
raw_data <- read.csv("C:/Users/deepa/Downloads/archive/METABRIC_RNA_Mutation.csv")
data <- raw_data

# Create binary outcome variable based on survival status
data$death_from_cancer <- factor(ifelse(data$overall_survival == 0, "YES", "NO"), levels = c("NO", "YES"))

# Create binary mutation label for TP53
data$tp53_mut_binary <- factor(ifelse(data$tp53_mut != "0" & data$tp53_mut != 0, "Mutated", "Not Mutated"))

# Convert treatments to factors
data <- data %>% mutate(
  chemotherapy = factor(ifelse(chemotherapy == 1, "Received", "Not Received")),
  hormone_therapy = factor(ifelse(hormone_therapy == 1, "Received", "Not Received")),
  radio_therapy = factor(ifelse(radio_therapy == 1, "Received", "Not Received"))
)

# Remove columns with >30% missing values
threshold <- 0.3 * nrow(data)
data <- data[, colSums(is.na(data)) < threshold]

# Impute missing values
getmode <- function(v) unique(v[which.max(tabulate(match(v, v)))])
data <- data %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%
  mutate(across(where(is.character), ~ factor(ifelse(is.na(.), getmode(.), .))))


# 03. PCA ON GENE EXPRESSION 
# Perform PCA to reduce dimensionality of gene expression data
expr_cols <- setdiff(names(data), c("patient_id", "overall_survival", "overall_survival_months", "death_from_cancer"))
expr_data <- data[, expr_cols] %>% select(where(is.numeric))
expr_data_scaled <- scale(expr_data)
pca_res <- PCA(as.data.frame(expr_data_scaled), graph = FALSE)
pca_scores <- as.data.frame(pca_res$ind$coord[, 1:5])
colnames(pca_scores) <- paste0("PC", 1:5)
data <- bind_cols(data, pca_scores)

```


```

